---
title: "585-Final-rmd"
author: "Ryan O'Dea"
date: "12/22/2020"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
pacman::p_load(
  ggplot2,
  tidyverse,
  forecast,
  tseries,
  timeDate,
  astsa
)
```

# Data Intake  

Taken from the [Federal Reserve Economics Data Database](https://fred.stlouisfed.org/series/TOTBKCR), the data chosen was the total bank credit (TBC) of all commercial banks. The TBC is the total amount of credit available to individuals or buisnesses from a banking institution and was measured in billions of USD. The  Data was taken weekly at the end of every Wednesday, from January 3, 1973 to April 1, 2020.  
The data had no noticeable outliers and contained every weekly data point. Two noticeable events during presented in the data were the 2009 banking crisis and the current COVID-19 pandemic. To garner a look at TBC without these two breakpoint events, a window between January 2012 and December 2019 was created and used henceforth.

```{r data-intake}
#Raw intake
cred <- read.csv("TOTBKCR.csv")

#TS coercion and subsetting
credts <- ts(cred$TOTBKCR, frequency = 52, start=c(1973,1))
window <- window(credts, 2012, 2019, 52)

 
```

```{r generic-plotting}
window %>% 
  autoplot() + 
  labs(title = "2012 - 2019 Total Banking Credit Availability",
       subtitle = "Data Source: https://fred.stlouisfed.org/series/TOTBKCR",
       ylab = "Billions USD")
```
The data was differenced twice to remove both the seasonality and linear trend components. Plotting the ACF and the PACF of the data yielded a PACF that decayed to zero and an ACF that appears to drop off after lag two or decay to zero. This makes possible models: ARIMA(0,2,2) or a low order, twice differenced ARIMA model. Utilizing the auto.arima function in R resulted in an ARIMA(2,2,2) model with $\phi_1= -.9408$ and $\phi_2=-.0423$ ; $\theta_1= -.0768$ and $\theta_2=-.8328$ or $x_t+.9408x_{t-1}+.0423x_{t-2}=e_t-.0768e_{t-1}-.8238e_{t-2}$

```{r decomposition}
#There appears to be a linear and seasonal trend, differencing twice is probable
decomp_win <- decompose(window)
decomp_win %>% 
  plot()
```

```{r differencing-data}
twice_dif <- diff(diff(window))
twice_dif %>%
  autoplot() +
  labs(title = "2012 - 2019 Twice Differenced Total Banking Credit Availability",
       subtitle = "Data Source: https://fred.stlouisfed.org/series/TOTBKCR")

twice_dif %>%
  ggAcf() + 
  labs(title = "Twice Differenced TBC Availability ACF")

twice_dif %>%
  ggPacf() + 
  labs(title = "Twice Differenced TBC Availability ACF")

#PACF appears to decay to zero, ACF roughly drops after lag one, but also shows non exponential decay with some lags spiking over the bounds.
#Possible a low order SARIMA model
```

```{r arima-processing} 
auto.arima(window)
#Appears to be a SARIMA(2,2,2) X (1,0,0) process 
sarima(window, 2, 2, 2, 1, 0, 0, 12)


arima_fit <- Arima(window, order = c(2, 2, 2), seasonal = c(1, 0, 0))
arima_fcast <- forecast(arima_fit, h =52)

arima_fcast %>%
  autoplot() + 
  labs(title = "TBC Forecast from SARIMA (2,2,2)x(1,0,0)[52]",
       subtitle = "Data Source: https://fred.stlouisfed.org/series/TOTBKCR") + 
  ylab("Billions USD")
```


```{r HW-forecasting}
#Holt-Winters forecast
hw_fit <- HoltWinters(window, seasonal = "multiplicative")
hw_fcast <- forecast(hw_fit, h=52)

hw_fcast %>%
  autoplot() + 
  labs(title = "TBC: HoltWinters Forecast",
       subtitle = "Data Source: https://fred.stlouisfed.org/series/TOTBKCR") + 
  ylab("Billions USD")
```


